Domain-specific working memory loads selectively increase negative interpertations of surprised facial expressions

### 2019-10-08 Manuscript/Manuscript.Rmd
update intro

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
## Working memory and load theory
Despite extensive research on the interaction of working memory and affective processes, there is still debate in the literature concerning how this cognitive process and emotional processes affect one another. Directly comparing working memory and self-regulation of emotional responses, Schmeichel and colleagues -[@schmeichel_working_2008] reported that individuals with higher levels of working memory capacity demonstrated improved self-regulation towards the emotional stimuli. This suggests a connection--perhaps through some shared resource pool--between mitigated emotional responding and larger working memory resource availablility. Other work has focused on the effects of moods or affective states on working memory performance. For instance, some reports claim that both positive and negative mood interfere with working memory (Eyesenck and Calvo, 1992); however, others suggest benefits of positive mood on working memory [@yang_positive_2012]. Similarly, working memory processes may alter concurrent affective processes. For instance, actively engaging working memory can mitigate emotional responses. Recent neuroimaging work reports that negative emotional responses decrease as the cognitive demands of a working memory task increase [@dillen_tuning_2008]. Additionally, depending on one's levels of trait-rumination, low trait-ruminators are better able to return heart rate and blood pressure to baseline levels when provided with a distractor task following an anger induction [@gerin_role_2006]. Together, these studies suggest a resource competition between cognitive and emotional processing; that is, when cognitive load demands are high (e.g., during active working memory maintenance), there are fewer resources available for other (i.e., affective) processes.

While previous work clearly suggests an inhibitory role for cognitive demands on emotional responses, researchers have primarily focused on emotional responses to clearly valenced emotional stimuli. For instance, Schmeichel and colleagues -[@schmeichel_working_2008] showed participants videos intended to elicit strong negative responses (e.g., disgust) or positive responses (e.g., humor), while others have focused on comparing responses to neutral and negative sitmuli [@dillen_tuning_2008], or even reducing symptoms of disorders with affective symptoms, such as anxiety [@spangler_little_2017]. However, many emotional appraisals in day-to-day life are more nuanced than those invoked by many of the types of negative stimuli one might encounter in the lab (e.g., snakes, mutilated bodies). For example, one may appraise the content of a billboard displaying a large order of french fries as either negative or positive depending on whether or not consuming that food would be (in)congruent with one's current goals. This emotional appraisal is completed under concurrent laod demands--that is, the perceiver must process both the emotional stimulus (i.e., the fries) as well as actively maintain their current goal state. If maintaining a goal state or some stimulus in working memory results in a large cognitive (not perceptual) load, then  Lavie's -[@lavie_load_2004] load theory posits that less executive resources are available to regulate incoming perceptual information. In turn, this could mean that more readily processed perceptual input (e.g., low spatial frequency images; LSF) might drive decisions under working memory loads.

Emotional stimuli readily capture attention compared to neutral stimuli, even in participants with amygdala damage [@hodsol_attentional_2011; @piech_attentional_2011]. Given emotional stimuli's priority position in the information processing stream, it may be that cognitive loads with emotional content, compared to neutral, differentially affect concurrent emotional appraisals. Indeed, negative emotional content slows performance on an n-back task [@kensinger_effect_2003]. Other domain-specific effects have been observed in many lines of research, including those beyond the working memory domain. For example, the Stroop task [@stroop_studies_1935], a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. Given the behavioral and neurological evidence for dissociations of emotional and non-emotional information domains in executive functions (e.g., working memory, inhibitory control), the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent emotional judgments.

<!-- Frequently, tasks in daily life require recruitment of working memory, but active working memory demands may affect concurrent processing. Many types of information can be held in working memory: strings of numbers, images, stimulus characteristics like brightness, spatial locations, and more. There is debate in the literature around the substrates of working memory and how this seemingly unitary construct is distributed across domain-specific subsystems (see [@baddeley_working_1986; @shah_separability_1996]). Recent work suggests though that working memory may contain more domain-specific subsystems than previously believed (e.g., affective working memory; e.g., @mikels_affective_2019). Similar to the work of Shah and colleagues -[@baddeley_working_1986], Mikels (find acutal study) showed ?double dissociation? of the affective working memory system from a more visually based working memory (i.e., maintenance of image brightness in memory). Similar domain-specific effects have been observed in many other lines of research, including those beyond the working memory domain. For example, the Stroop task [@stroop_studies_1935], a common measurement tool for inhibitory control, has been modified to include both emotional and non-emotional (neutral) stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Neuroimaging work also supports the notion that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008], suggesting a division of subsystems. Given the behavioral and neurological evidence for dissociations of information domains in working memory, task interference and more, the present work aims to clarify the interaction of emotional and non-emotional visual working memory demands on concurrent emotional judgments (i.e., interpretations of emotional facial expressions). -->

<!-- While it is well known that distractors can undermine task performance [@hagen_effect_1967; cite, cite), some evidence suggests a role for domain-specificity in task interference.  For instance, domain-specific interference further exacerbates these task effects compared to domain-general stimuli [@gruber_effects_2001]. Within an affect-specific context, the literature reports conflicting results for the role of affective stimuli on working memory processing (see @jackson_enhanced_2009 and @mather_emotional_2006), with some suggesting ehnaced recall for emotional stimuli and others reporting impairment in working memory processes. The relationship between these systems is bidirectional though, and often times working memory affects affective processing. One such way is through dampening emotional responses, and converging work suggests that cognitive demands can mitigate negative emotional effects [@dillen_tuning_2008]. One explanation for mitigated emotional responding during concurrent working memory demands is load theory. This theory posits that cognitive resources are finite, thus limiting the availability for simultaneous processing under large task demands [@lavie_load_2004]. One limitation to this previous work is that the affective processes under study involved clearly interpretable stimuli/feelings. A more malleable affective process, such as interpreting ambiguously valenced emotional facial expressions, may offer better understanding of how concurrent working memory demands affect emotional processes. -->

## The contribution of valence bias
<!-- Facial expressions are rich with affective information, and correctly interpreting these social cues is critical for successfully navigating the social world.--> As mentioned above, individuals differ in their tendency to interpret ambiguously valenced stimuli, like a tempting food item, as either positive or negative. <!-- Additionally, facial expressions can serve as ambiguously valenced stimuli. While most facial expressions send a clear social signal, this is not always the case. While a smile likely expresses a positive affective state, other cues are not so clear.--> This tendency is known as one's *valence bias*, and a growing body of work has used both facial expressions and scenes (or items) to quantify this individual difference (CITE). Biases towards negativity are related to psychopathology, such as depression and anxiety [@matthews_cognitive_2005]

<!--For instance, a surprised expression could signal either a positive (e.g., winning the lottery) or negative (e.g., seeing a snake in the woods) affective state in the expresser. When contextual information is limited, individuals differ in their tendency to interpret surprised facial expressions as positive or negative. Importantly, this affective bias extends beyond facial expressions, as individuals often show a similar bias to surprised faces as they do for ambiguous scenes [@neta_neural_2013] or even words (Harp, Petro, Brown & Neta, in prep). This bias towards positive or negative interpretations is known as one's valence bias and myriad factors contribute to one's bias. -->

Both bottom-up (e.g., perceptual input) and top-down (e.g., stereotypes) processes are recruited for interpretation of facial expressions, and a growing body of work suggests that the initial interpretation of emotionally ambiguous stimuli is negative and driven by bottom-up processes. For instance, reaction times are faster for negative interpretations of ambiguous stimuli (just faces??) [@neta_dont_2016]. Additionally, presentation of surprised facial expressions as low spatial frequency images, which is processed more readily than high spatial frequency images, biased interpretations towards negativity [@neta_primacy_2010]. Consequently, under this framework, arriving at a positive interpretation requires additional, top-down regulatory processes and there is evidence to support this as well. For example, forcing participants to slow their responding during interpretations of ambiguous images shifts individuals' biases towards positivity [@neta_its_2018]. Neuroimaging evidence supports this initial negativity hypothesis as well, more positive individuals show higher levels of BOLD activation in brain regions recruited during emotion regulation [@petro_individual_2018]. <!-- Perceptual input also contributes to valence bias. In one recent study, Neta and colleagues [-@neta_all_2017] showed that faster intial fixation, as well as longer overall fixation, on the mouth is related to more positive interpretations of surprised faces and that forcing gaze patterns to match those of modulated interpretations of surprised expressions. THIS NEEDS TO BE REWORKED --> In short, slowing response time allows individuals to better regulate the potentially negative information in surprised expressions and to see it in a more positive light.

Given that a regulatory mechanism likely contributes to positive interpretations of surprised facial expressions, domain-specific interference may cause more negative interpretations of ambiguity compared to a more domain-general interference. Mattek and colleagues [-@mattek_differential_2016] recently showed that different levels of cognitive load (i.e., holding either a single or seven digit number in working memory) does not affect subjective interpretations of surprised facial expressions, but that high cognitive loads do mitigate mouse trajectories. While the authors interpret this as a distinction between trait-like biases and dynamic cognitive-motor processes, there may be more domain-specific processes (e.g., emotional components) that span across these two measures of valence bias. Given the task irrelevance of the numeric distractors in Mattek and colleagues' [-@mattek_differential_2016] work, it follows that the resources required for interpreting ambiguity as positive [@neta_corrugator_2009] may not have been recruited for working memory maintenance, and thus no change in subjective ratings was observed.

## The present study
In the present study, we aim to test the effects of low and high working memory loads in both emotional and neutral domains. We expect that trials in which participants are maintaining an emotional working memory load will be more negative than neutral trials. Further, we predict that higher working memory laod trials, specifically in the emotional domain, will result in even more exaggerated negative interpretations.

<!-- Despite the trait-like nature of this bias [@neta_corrugator_2009], valence bias may be shifted, at least temporarily, by a number of experimental manipulations. As mentioned above, simple manipulations like slowing response times will shift bias [@neta_its_2018]. Additional work has shown that increases in salivary cortisol after a stressor relate to more negative interpretations of surprised faces from baseline to post-stressor [@brown_cortisol_2017]. Similarly, participants with positive biases at baseline will interpret surprise as more negative under threat of shock [@neta_impact_2017], suggesting that taxing cognitive resources, in this case attention, reduces the ability of individuals to interpret surprised faces as positive. -->

Recent work suggests that ambiguity resolution in this context requires more cognitive resources/processing compared to clearly valenced faces [@mattek_differential_2016; @neta_dont_2016]. 
The valence bias is trait-like [@neta_corrugator_2009] and generalizes to non-face stimuli [@neta_neural_2013]; however, it is also malleable and may differ depenending on experimental manipulations, including stress inductions or instructions to slow responding [@brown_cortisol_2017; @neta_dont_2016]. Importantly, the valence bias relates to behavior outside of the laboratory; specifically, it is known to relate to depressive symptomology [@petro_initial_2018], at least in children. Chronic negativity biases are common in numerous psychopathologies, including depression and anxiety [@matthews_cognitive_2005].

<!-- Distractors and task irrelevant stimuli often have detrimental effects on performance in a variety of tasks [; cite, cite]. Further, domain-specific interference may further exacerbate these effects compared to domain-general stimuli [@gruber_effects_2001]. This effect holds up in the emotional domain; for example, the Stroop task [@stroop_studies_1935] has been modified by some researchers to include emotional stimuli [@whalen_emotional_2006] which has pronounced effects when the emotional words are population specific (e.g., trauma words in a PTSD sample). Indeed, neuroimaging work supports the idea that separate systems handle attentional biasing for domain-specific (emotional vs. non-emotional) task relevancy [@egner_dissociable_2008]. -->

<!-- Notes: Allie showed that depletion doesn't affect ratings, but we thought that there may be an effect of specific domains of depletion (i.e., neutral vs. emotional) on ratings. Focus on EMO > NEU ratings and POS > NEG MDs. MT analyses could be supplementary material, but you'll be going down the rabbit hole.  -->
<!-- Recently, others have suggested a dynamic social trait space model for judgments of faces [@stolier_dynamic_2018], and it may be important to consider additional components of the categorization process, such as domain-specific loads.  -->

>>>>

### load necessary libraries ###
suppressPackageStartupMessages(library(readxl)) 
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(openxlsx))
suppressPackageStartupMessages(library(Rmisc))
suppressPackageStartupMessages(library(broom))
```
# Introduction
## Facial expressions and individual differences

### Facial expressions are an important social signal
Facial expressions are important social signals, often communicating emotion between individuals and even sparking emotional responses in the perceiver [@frith_role_2009]. Researchers have long argued that facial expressions are universal signals of emotion, positing humans' innate ability to associate facial expressions with emotional experience [@darwin_expression_1872; @ekman_constants_1971; @izard_innate_1994]. More recently, Barrett and colleagues suggested that the relationship between facial expressions and emotions is more complex, stating that expression of an emotion may vary across cultures, social context, or even individuals -[@barrett_emotional_2019]. Other work has shown individual differences in interpretations of facial expressions [@green_factors_2018; @neta_corrugator_2009]. Regardless of this variability in emotional expressions, experience, and interpretations, humans readily make judgments about personality traits (e.g., trustworthiness), aesthetics (e.g., attractiveness), and emotions from faces [@todorov_evaluating_2008; @said_statistical_2011; @carroll_facil_1996]. Interpretations of valence, that is the inherent positive or negative emotional value of a stimulus, are one instance of judgments of facial expressions guiding potential social (i.e., approach-avoidance) behavior <!-- not sure if this is truly social? --> [@krieglmeyer_being_2010].

### Individuals differ (trait) in their tendency to interpret ambiguous images, like a surprised face
While most people can accurately differentiate the emotional valence of many facial expressions [CITE], such as consistently interpreting angry (happy) faces as negative (positive), there are individual differences in valence judgments of emotionally ambiguous facial expressions, like a surprised face [@neta_corrugator_2009; @petro_individual_2018 <!-- right citations?-->]. This difference in valence interpretations of surprised expressions is attributable to such stimuli's predictive value for both positive and negative outcomes in an individual's previous experience. For instance, a surprised expression could signal positive (e.g., winning the lottery) or negative (e.g., a car accident) events. These differences in valence interpretations represent an important individual difference, as the same stimulus can result in two alternative interpretations between individuals--likely leading to different downstream behaviors [e.g., @krieglmeyer_being_2010]. This individual difference in interpretations of emotionally ambiguous stimuli is known as one's *valence bias*, and a growing body of work has used both facial expressions and scenes to better understand this individual difference [@neta_corrugator_2009; @neta_primacy_2010; @neta_neural_2013].

### Despite the bias, everyone seems to show an initial negativity--positivity is associated with emotion regulation
Despite one's valence bias, the initial response to ambiguity appears to be negativity [@neta_corrugator_2009; @neta_primacy_2010; @neta_valence_2011; @petro_individual_2018]. Under this framework, which is known as the *initial negativity* hypothesis, positive interpretations rely on the implementation of some emotion regulation strategy, perhaps similar to cognitive reappraisal. During cognitive reappraisal, individuals work to change intial perceptions of an emotional stimulus [@lazarus_short_1964]. Similarly, the initial negativity hypothesis posits that individuals' initial perception of surprised expressions is negative, and that those arriving at a positive interpretation must implement a regulatory mechanism to alter their interpretation.

### Behavioral and brain data supports the initial negativity hypothesis
Both behavioral and brain data in support of the initial negativity hypothesis come from many studies. For instance, reaction time (RT) data show that individuals with more positive biases take longer to reach a valence judgment for surprised expressions than those with a more negative bias [@neta_corrugator_2009], suggesting a more time-intensive (regulatory) process for positive interpretations. Other work demonstrates the automaticity of negative interpretations through presenting either low- (LSF) or high-spatial frequency (HSF) images of surprised expressions. The images with only LSF information, which is processed earlier than its HSF counterpart, were rated more negatively than the HSF images [@neta_primacy_2010]. Additionally, surprised facial expressions are more quickly detected among happy (positive) than angry (negative) faces [@neta_valence_2011]. The results from this emotional oddball paradigm suggest that surprised expressions are more readily perceived as similar to angry faces (i.e., perceived as negative) than happy faces. There is also support for the initial negativity hypothesis in studies with brain data. Ventromedial prefrontal cortex, a putative regulatory region, and the amygdala actively are inversely correlated, and participants with more negative biases show higher activity in the amygdala while more positive participants show higher activity in vmPFC [@kim_inverse_2003]. More recently, Petro and colleagues showed that participants with a more positive valence bias show more activity for surprised faces in emotion regulation-related brain regions -[vmPFC; @petro_individual_2018].Taken together, these effects support the idea that negative interpretations of surprised faces are faster and default process, while positive interpretations rely on slower regulatory processes. <!-- is it okay to combine these sections? -->

## Cognitive loads and task interference
### In our daily lives, these resources are not always available

### Active WM maintenance affects concurrent processes

### Other work has tested the effects of cognitive load on valence bias, but did not find an effect for interpretations of surprise. This could be due to domain-specific effects of cognitive load.

## The present study

### h1: domain main effect

### h2: domain x load interaction
